---
format: html
server: shiny
---

## Chapter 2: Working with Descriptive Statistics

Descriptive statistics *describe* what the data *looks* like. You likely recall seeing the bell curve in PSY248 (statistics). You may even remember the phrase *central tendency*. The idea behind these values is that they give us a metric of where the data (likely) comes from.

Let's go over the usual suspects: Mean, Median, Mode, and Range

#### Mean

The mean represents the average of a set of numbers. Mathematically represented by the following equation:

$\overline{X}=\frac{\Sigma x}{n}$

In a simple group, the mean can be seen by *eyeballing* the data:

```{r}
library(huxtable)
library(dplyr)
library(ggplot2)
library(shiny)

x <- rnorm(5,1,0)
x
```

The *only* number that exists here is 1, so the mean *has* to be 1...right?!

```{r}
paste0("The mean of x = ", mean(x))
```

However, when the numbers become more spread out and varied, it is not as easy.

```{r}
y <- rnorm(15,5,2) |> round(0)
y
```

Eyeballing this data will not do us any good!

```{r}
paste0("The mean of y = ", mean(y) |> round(2))
```

$\sigma = \frac{\Sigma (x- \overline{x} }{n-1}$

::: callout-tip
## Never trust the eyeballing of data as a replacement for calculation!
:::

#### Test

```{r}
#| context: server
library(shiny)

ui <- fluidPage(
  textInput("numba","Enter N",value = ""),
  plotOutput("plotme")
)

server <- function(input, output, session) {
  output$plotme <- renderPlot({
    x <- rnorm(input$numba
               ,100,15)
    plot(x)
  })
}

shinyApp(ui, server)
```
